---
title: "Regression"
author: "Nirzaree"
date: "31/08/2020"
output: html_document
---

Goal: Understand regression concepts with examples

* What is regression?   
  A technique applied to estimate a continuous output variable as a function of one or more input variables.

* What types of regression techniques exist? 

1. Linear: $$y = a_0 + a_1*x_1 + a_2*x_2 + a_3*x_3 + \epsilon$$    
  + Univariate  
  + Multivariate  
2. Polynomial: $$y = a_0 + a_1*x_1 + a_2*(x_1)^2 + a_3*(x_2) + a_4*(x_2)^2 + a_5*x_1*x_2 + \epsilon$$
3. Logistic: $$log(p_i/(1-p_i)) = a_0 + a_1*x_1 + a_2*x_2 + a_3*x_3 + \epsilon$$

##Datasets {.tabset}

### Boston Housing 

```{r housingsetup,include=FALSE,echo = FALSE, message = FALSE, warning = FALSE,fig.width = 16, fig.height = 10}
library(data.table)
library(mlbench)
library(corr)
```

About the features: 
CRIM - per capita crime rate by town : Should have an impact (negative correlation)
ZN - proportion of residential land zoned for lots over 25,000 sq.ft. (?)
INDUS - proportion of non-retail business acres per town. (?)
CHAS - Charles River dummy variable (1 if tract bounds river; 0 otherwise) (maybe positive correlation)
NOX - nitric oxides concentration (parts per 10 million) (negative correlation)
RM - average number of rooms per dwelling (positive correlation)
AGE - proportion of owner-occupied units built prior to 1940 (?)
DIS - weighted distances to five Boston employment centres (negative correlation)
RAD - index of accessibility to radial highways (positive correlation)
TAX - full-value property-tax rate per $10,000 (?)
PTRATIO - pupil-teacher ratio by town (negative correlation)
B - 1000(Bk - 0.63)^2 where Bk is the proportion of blacks by town (negative correlation)
LSTAT - % lower status of the population (negative correlation)
MEDV - Median value of owner-occupied homes in $1000's

```{r housingload,include=FALSE,echo = FALSE, message = FALSE, warning = FALSE,fig.width = 16, fig.height = 10}
data("BostonHousing2")
summary(BostonHousing2)
```

```{r FeatureSelection,echo = FALSE, message = FALSE, warning = FALSE,fig.width = 16, fig.height = 10}
plot(x = BostonHousing2$crim,y = BostonHousing2$medv) #negative effect 
plot(x = BostonHousing2$rm,y = BostonHousing2$medv)
plot(x = BostonHousing2$nox,y = BostonHousing2$medv)
plot(x = BostonHousing2$lstat,y = BostonHousing2$medv)
plot(x = BostonHousing2$tax,y = BostonHousing2$medv)

cor(BostonHousing2$rm,BostonHousing2$medv)

RoomsModel <- lm(medv~rm,data = BostonHousing2)
summary(RoomsModel) #not bad 

FewVarModel <- lm(medv~rm+crim+lstat+ptratio,BostonHousing2)
summary(FewVarModel)

dtBostonHousing <- data.table(BostonHousing2)
dtBostonHousing[,"cmedv" := NULL]

AllVarModel <- lm(medv~crim+zn+indus+chas+nox+rm+age+dis+rad+tax+ptratio+b+lstat,dtBostonHousing)
summary(AllVarModel)

plot(AllVarModel)
```

