---
title: "Regression with Keras"
author: "Nirzaree"
date: "23/09/2020"
output: html_document
---
https://tensorflow.rstudio.com/tutorials/beginners/basic-ml/tutorial_basic_regression/
```{r setup,include=FALSE,echo = FALSE, message = FALSE, warning = FALSE,fig.width = 16, fig.height = 10}
library(keras)
library(tfdatasets)
library(tidyverse)
library(ggplot2)
```

```{r loadData,include=FALSE,echo = FALSE, message = FALSE, warning = FALSE,fig.width = 16, fig.height = 10}
boston_housing <- dataset_boston_housing()

c(train_data,train_labels) %<-% boston_housing$train
c(test_data,test_labels) %<-% boston_housing$test
```

```{r inspectdata,include=FALSE,echo = FALSE, message = FALSE, warning = FALSE,fig.width = 16, fig.height = 10}
train_data[1,]

#add column names
column_names <- c('CRIM', 'ZN', 'INDUS', 'CHAS', 'NOX', 'RM', 'AGE', 
                  'DIS', 'RAD', 'TAX', 'PTRATIO', 'B', 'LSTAT')

#some fancy dplyr + tibble stuff in the tutorial which I am ignoring for now.
#cant ignore because see note below
train_df <- train_data %>% as_tibble(.name_repair = "minimal") %>% setNames(column_names) %>% mutate(labels = train_labels)

test_df <- test_data %>% as_tibble(.name_repair = "minimal") %>% setNames(column_names) %>% mutate(labels = test_labels)
```

todo: understand why standard scaling and not simply normalizing 
Also in normal regression using this dataset, no preproc is done

```{r Preprocess,include=FALSE,echo = FALSE, message = FALSE, warning = FALSE,fig.width = 16, fig.height = 10}
spec <- feature_spec(train_df,labels ~ .) %>% step_numeric_column(all_numeric(),normalizer_fn = scaler_standard()) %>% fit() 
#note : so the above didnt work without doing all the fancy dplyr tibble stuff. It was to create a dataset that is acceptable to tensorflow

spec
```

```{r Model,include=FALSE,echo = FALSE, message = FALSE, warning = FALSE,fig.width = 16, fig.height = 10}
build_model <- function() {
  input <- layer_input_from_dataset(train_df %>% select(-labels))

  output <- input %>% layer_dense_features(dense_features(spec)) %>% 
    layer_dense(units = 64,activation = 'relu') %>% 
    layer_dense(units = 64,activation = 'relu') %>%
    layer_dense(units = 1)

  model <- keras_model(input,output)

  summary(model)

  model %>% compile(
    loss = "mse",
    optimizer = optimizer_rmsprop(),
    metrics = list("mean_absolute_error")
  )
  model
}
```

```{r ModelBuild,include=FALSE,echo = FALSE, message = FALSE, warning = FALSE,fig.width = 16, fig.height = 10}
print_dot_callback <- callback_lambda(on_epoch_end = function(epoch,logs) {
  if (epoch %% 80 == 0) cat("\n")
  cat(".")
})

model <- build_model()

history <- model %>% fit(
  x = train_df %>% select(-labels),
  y = train_df$labels,
  epochs = 500,
  validation_split = 0.2,
  verbose = 0,
  callbacks = list(print_dot_callback)
)
```
```{r CheckModel,include=FALSE,echo = FALSE, message = FALSE, warning = FALSE,fig.width = 16, fig.height = 10}
plot(history)
```

Little improvement after certain epochs (> 200). Let's update the fit to stop running when there's little improvement. 
```{r preventoverfitting,include=FALSE,echo = FALSE, message = FALSE, warning = FALSE,fig.width = 16, fig.height = 10}
early_stopping <- callback_early_stopping(monitor = "val_loss",patience = 20)

model <- build_model()

history <- model %>% fit(
  x = train_df %>% select(-labels),
  y = train_df$labels,
  epochs = 500,
  validation_split = 0.2,
  verbose = 0,
  callbacks = list(early_stopping)
)

plot(history) #fix this: todo: error: arguments imply differing number of rows: 500, 260, 2000
```

```{r Validateontestset,include=FALSE,echo = FALSE, message = FALSE, warning = FALSE,fig.width = 16, fig.height = 10}
c(loss, mae) %<-% (model %>% evaluate(test_df %>% select(-labels), test_df$labels, verbose = 0))
paste0("MAE on test set $",sprintf("%.2f",mae*1000))
```

```{r PredictOnTestData,include=FALSE,echo = FALSE, message = FALSE, warning = FALSE,fig.width = 16, fig.height = 10}
test_predictions <- model %>% predict(test_df %>% select(-labels))

test_predictions
```

Summary:
Too much new info here. 

1. Some crazy conversion to tensorflow dataframe
2. Preprocess : some crazy functions to do standardscaling for all features
3. Model function: 
4. Model fit with history 
5. Plot history
6. Early stopping for preventing overfitting
7. Model validation using MAE 
8. Prediction on new data