---
title: "Time Series Analysis"
author: "Nirzaree"
date: "23/10/2020"
output: 
  html_document:
    fig_caption: yes
    toc: yes
    toc_collapsed: yes
    toc_float: yes
toc_depth: 3
---

## Theory
```
Stationary timeseries:
1. Mean of series is constant and does not vary with time. 
2. Variance is constant (homoscedasticity)
3. Covariance is constant (seasonality effect is minimal)
=> Should look like random white noise irrespective of observed time interval. 

Stationary testing and converting a timeseries into a stationary series is crucial for any time series analysis. 

Ref: 
https://www.analyticsvidhya.com/blog/2015/12/complete-tutorial-time-series-modeling/
http://r-statistics.co/Time-Series-Analysis-With-R.html
https://boostedml.com/2020/05/visualizing-time-series-in-r.html
http://people.cs.pitt.edu/~milos/courses/cs3750/lectures/class16.pdf  
```
## AirPassengers
### Explore dataset
```{r setup,echo = FALSE, message = FALSE, warning = FALSE,fig.width = 16, fig.height = 10}
library(ggfortify)
library(zoo)
library(tseries)
library(astsa)
library(forecast)
library(ggplot2)
```


```{r ExplorationofTimeseries,echo = TRUE, message = FALSE, warning = FALSE,fig.width = 16, fig.height = 10}
data("AirPassengers")
class(AirPassengers)

start(AirPassengers)
end(AirPassengers)

frequency(AirPassengers)
summary(AirPassengers)

plot(AirPassengers) + abline(reg = lm(AirPassengers~time(AirPassengers)))

cycle(AirPassengers)

plot(aggregate(AirPassengers,FUN = mean))

boxplot(AirPassengers~cycle(AirPassengers))
```

**Quick Observations:**  

* Trend: Increasing
* Seasonality: Repeated patterns over the years.
* Increasing variance: Swings are increasing over the years. 

A time series has 4 components:

1. Trend: Defines the long term behavior
2. Cycles: Define mid term non repeated deviations from trend
3. Seasonality: Defines periodic or repeated fluctuations
4. Noise or remainder: Random fluctuations

In many cases, trend and cycles are combined into one single trend-cycle or trend component.
The decomposition of time series is either additive or multiplicative 

$$ x_t = T_t + S_t + R_t $$ or 
$$ x_t = T_t * S_t * R_t $$

When to use which formulation: 
If magnitude of seasonality depends on magnitude of raw values, then multiplicative else additive.
In the given dataset, seasonality does depend on the raw values so multiplicative decomposition is more relevant.

```{r DecomposedTS,echo = TRUE, message = FALSE, warning = FALSE,fig.width = 16, fig.height = 10}
decomposedAP <- decompose(AirPassengers,type = "multiplicative")
autoplot(decomposedAP)

additivedecomposedAP <- decompose(AirPassengers,type = "additive")
autoplot(additivedecomposedAP)
```

**Observations:**

* Trend component is almost linear
* Each year starts with a low, with peak in July - August summer followed by a dip again.  
* Random component is not completely random (although I dont see any real pattern, but the [reference](https://boostedml.com/2020/05/visualizing-time-series-in-r.html) says so) (todo: understand this better)

### Stationarity test
* **Augmented Dickey Fuller Test (ADF)**: Most common test to detect stationarity. Here, the null hypothesis is that the series is non-stationary. So if we find that the p-value in the ADF test < significance level (0.05), we reject the null hypothesis.  
```{r stationaritytest,echo = TRUE, message = FALSE, warning = FALSE,fig.width = 16, fig.height = 10}
adf.test(AirPassengers)
```
**Observation:** Since the p-value is not less than 0.05, we conclude that the time-series is not stationary.

* **KPSS** test: Test for trend stationarity.

```{r trendstationarity,echo = TRUE, message = FALSE, warning = FALSE,fig.width = 16, fig.height = 10}
kpss.test(AirPassengers)
```
* **PP Test**:

### Stationarize the series:

* De-trend:


* De-seasonalize: 
```{r detrend,echo = TRUE, message = FALSE, warning = FALSE,fig.width = 16, fig.height = 10}
plot(log(AirPassengers))

plot(diff(log(AirPassengers)))

adf.test(diff(log(AirPassengers)))
```

**Observations:** looks stationary but the p-value doesn't say so as much. 

### Model

### Predict:

Ref and issues:

1. http://people.cs.pitt.edu/~milos/courses/cs3750/lectures/class16.pdf theory only
2. http://r-statistics.co/Time-Series-Analysis-With-R.html No model. no good transformation steps
and explanations.
3. https://www.kaggle.com/chirag19/time-series-analysis-with-python-beginner has everything, but 
dont know how credible + python (not R)
4. https://rpubs.com/neharaut05/TimeSeries_AirPassangerForecast just log(ts) for stationarizing.  Not solid.
5. http://rstudio-pubs-static.s3.amazonaws.com/311446_08b00d63cc794e158b1f4763eb70d43a.html No transformation of the series at all
6. https://www.analyticsvidhya.com/blog/2015/12/complete-tutorial-time-series-modeling/ too much of things without really basing the concepts. 3


https://www.kaggle.com/chirag19/air-passengers/notebooks
https://www.kaggle.com/chirag19/time-series-analysis-with-python-beginner

https://rpubs.com/neharaut05/TimeSeries_AirPassangerForecast
http://rstudio-pubs-static.s3.amazonaws.com/311446_08b00d63cc794e158b1f4763eb70d43a.html