---
title: "Iris"
author: "Nirzaree"
date: "10/12/2020"
output: 
  html_document:
    toc: true
    toc_float: true
---
### 1. Setup

```{r 1_setup, echo=TRUE}
library(data.table)
```

### 2. Load Data
```{r 2_loaddata,echo=TRUE}
dtIris <- as.data.table(iris)
```

### 3. Explore Data
```{r 3_exploredata,echo=TRUE,message=FALSE}
library(caret)
summary(dtIris)
sapply(dtIris, class)
summary(dtIris[,Species])
featurePlot(dtIris[,1:4],dtIris[,Species],"density")
```
**Observations:**  
1. Balanced Classes: Exactly same proportions of each of the 3 species of Iris.    
2. Scale & Range of each feature column: Same unit in each column and not a lot of variation among the scales and ranges.    

### 4. Data Preprocessing
```{r 4_datapreprocessing,echo=TRUE}
#split into train test sets. 
set.seed(100)
trainingindices <- sample(seq(1:nrow(dtIris)),0.8*nrow(dtIris))
dtTrainIris <- dtIris[trainingindices,]
dtTestIris <- dtIris[-trainingindices,]
```

### 5. Apply Models on training set

```{r 5_Models,echo=TRUE}
library(caret)
fClassificationModels <- function(modelFormula,dtTrain,trainConfig,evalMetrics,modelList) {
 AllModels <- lapply(modelList, function(x) {
    SpecificModel <- train(modelFormula,
                           data = dtTrain,
                           method = x,
                           trControl = trainConfig,
                           metric = evalMetrics)
    return(SpecificModel)
  })
  return(AllModels)
}

AllIrisModels <- fClassificationModels(modelFormula = Species~.,
                                       dtTrain = dtTrainIris,
                                       trainConfig = trainControl(method = "cv",number = 10),
                                       evalMetrics = "Accuracy",
                                       modelList = c("rpart","rf","nb","knn","lda","svmRadial"))
```

### 6. Find best model

```{r 6_CompareModels,echo=TRUE}
AllModelsSummary <- resamples(list(
  DecisionTree = AllIrisModels[[1]],
  RandomForest = AllIrisModels[[2]],
  NaiveBayes = AllIrisModels[[3]],
  knn = AllIrisModels[[4]],
  LDA = AllIrisModels[[5]],
  svmRadial = AllIrisModels[[6]]
))

dotplot(AllModelsSummary)
summary(AllModelsSummary)
```

### 7. Predict using best model

```{r 7_Predict,echo=TRUE}
#LDA looks best
predictIris <- predict(AllIrisModels[[5]],
                           dtTestIris)

confusionMatrix(predictIris,dtTestIris[,Species])
```

### Summary
* What was not done in this project: Hyperparameter tuning, as even without it, the results are pretty good. 
* Not a lot of preprocessing: as the features are on the same scale and ranges. 
